---
title: "Fetal Health Classification"
author: "Aloui Ameni Roua Riahi"
date: "2023-05-19"
output: html_document
---


** Description of Dataset **

This dataset contains 2126 records of features extracted from Cardiotocogram exams, which were then classified by expert obstetrician into 3 classes: "Normal", "Suspect" and "Pathological". The features included in this study are the following:

baseline value: Baseline Fetal Heart Rate (FHR) (beats per minute)
accelerations: Number of accelerations per second
fetal_movement: Number of fetal movements per second
uterine_contractions: Number of uterine contractions per second
light_decelerations: Number of light decelerations (LDs) per second
severe_decelerations: Number of severe decelerations (SDs) per second
prolongued_decelerations: Number of prolonged decelerations (PDs) per second
abnormal_short_term_variability: Percentage of time with abnormal short term variability
mean_value_of_short_term_variability: Mean value of short term variability
percentage_of_time_with_abnormal_long_term_variability: Percentage of time with abnormal long term variability
mean_value_of_long_term_variability: Mean value of long term variability
histogram_width: Width of histogram made using all values from a record
histogram_min: Histogram minimum value
histogram_max: Histogram maximum value
histogram_number_of_peaks: Number of peaks in the exam histogram
histogram_number_of_zeroes: Number of zeros in the exam histogram
histogram_mode: Histogram mode
histogram_mean: Histogram mean
histogram_median: Histogram median
histogram_variance: Histogram variance
histogram_tendency: Histogram tendency
fetal_health: Encoded as 1-Normal; 2-Suspect; 3-Pathological Which is our target column.
```{r}
library(tidyverse)
library(kableExtra)
library(skimr)
library(corrplot)
library(e1071)
library(caret)
library(gbm)
library(MASS)
library(ggplot2)
```

```{r }
data=read.csv("C://Users//aloui//OneDrive//Bureau//Ameni//R//projetFoetus//Projet//fetal_health.csv",sep=',',dec='.')
View(data)
head(data)

```

** Data cleansing **

First of all we'll be checking if there are any missing values or categorical variables.
```{r}
str(data)
colSums(is.na(data))

```
All of the variables are numeric, we'll be transforming our target into factor.There is no missing value. 

```{r}
data$fetal_health=as.factor(data$fetal_health)
str(data$fetal_health)
```

** Exploratory Data Analysis **

```{r}
target <- 'fetal_health'
data[,target] <- as.factor(data[,target])
plot(data[,target])
grid()

```
Next comes visualization of the dataset features and their relations to the target fetal_health. The following  figure portray sets of feature boxplots by classes of fetal_health. In general, the boxplots reveal clear class-related differences in the distributions of the features.

```{r}
data_long <- gather(data, key = 'variable', value = 'value', -fetal_health)

# Créer le graphique
ggplot(data_long, aes(x = '', y = value, fill = fetal_health)) +
  facet_wrap(~ variable, scales = 'free') +
  geom_boxplot() +
  coord_flip() +
  labs(x = NULL, y = NULL) +
  theme(axis.text.x = element_text(angle = 45)) +
  scale_fill_manual(name = 'Fetal health',
                    guide = guide_legend(reverse = TRUE),
                    labels = c('Normal (1)', 'Suspect (2)', 'Pathological (3)'),
                    values = c('forestgreen', 'gold', 'red'))


```

Correlation

```{r}
library(corrgram)
library(corrplot)
x <- data[, sapply(data, is.numeric)]
corr=cor(x)
corrplot(corr, method="circle",tl.cex=0.6)
```

From figures, we can find “histogram_mean”, “histogram_mode” and “histogram_median” have high correlations value. We also notice a high correlation between these three and  “baseline.value”. Moreover, “histogram_width” has high correlation with “histogram_max” and “histogram_number_of_peaks”.



** Data Standardization **

Before strating we will do a feature standardization to helps adjust all data elements to a common scale in order to improve the performance of PCA and prediction.

```{r}
data_new=scale(data[,-22])
head(data_new,2)
data=data.frame(data_new,data[22])
View(data)
str(data)
```

** Principal Component Analysis(PCA) **

We removed the fetal_health column which is not necessary for the  PCA.
```{r warning=FALSE}
library(FactoMineR)
pca.data=PCA(data_new,graph = FALSE)
pca.data$eig
```

```{r}
library(factoextra)
fviz_screeplot(pca.data)

```
 We are going to choose the elbow criterion for the choice of the number of dimensions, the offset is between the 3rd and the 4th component, we will then retain the first 3 components.

* Study of variables *

```{r}
fviz_pca_var(pca.data,c(1,2))
fviz_pca_var(pca.data,c(2,3))
```

```{r}
pca.data$var$coord
round(pca.data$var$cos2,3)
round(pca.data$var$contrib,3)
```
Based on these results, the contributions of the components in each dimension are clear. dim1 consists of "histogram_width", "histogram_min" and ect. To visualize, we will also draw the contribution of the plot variables.


```{r}
fviz_contrib(pca.data,'var',axes = 1)
fviz_contrib(pca.data,'var',axes = 2)
fviz_contrib(pca.data,'var',axes = 3)
```
Study of individuals

```{r}
fviz_pca_ind(pca.data,col.ind = 'cos2',geom = 'point')
fviz_pca_ind(pca.data,geom.ind = "point", repel = TRUE,col.ind = "contrib")

```
* Visualisations pour la valeur de groupe (fetal_health) *

Cette fois, Nous allons visualisez non seulement les observations et les variables, mais nous allons différenciez Fetal_health (1-Normal, 2-Suspect, 3-Pathological).

```{r}

```

We find that:

Group 3, which is pathological, has important variables “severe decelerations” and “prolonged_decelerations”, which indicates fetal health problems.

group 2 which is suspected of being unhealthy has the variables "abnormal_short_term_variability", "percentage_of_time.." and histogram_min" which show signs of suspected health problems.

group 1 which is normal has larger residual variables.

then, we also see that the observations of group 3 (pathological) are more spread out and the variance rather greater than in the other groups.

Logistic Regression
```{r}
colnames(data)
```

```{r}
set.seed(123)
x2<-sample(c(1:nrow(data)), 0.8*nrow(data))
training=data[x2,]
training
test=data[-x2,]
library(nnet)
library(MASS)

# Fit the model
model <- nnet::multinom(fetal_health ~., data = training)
# Summarize the model
summary(model)
modele.backward <- stepAIC(model, ~. , trace = TRUE, data = training, direction = "backward")


```
the selcted model is the following: 
fetal_health ~ baseline.value + uterine_contractions + 
    prolongued_decelerations + abnormal_short_term_variability + 
    percentage_of_time_with_abnormal_long_term_variability + 
    mean_value_of_long_term_variability + histogram_width + histogram_min + 
    histogram_mode + histogram_mean+histogram_variance  
    +fetal_movement+histogram_number_of_peaks+accelerations 
prediction:

```{r warning=FALSE}
library(magrittr)
test2=test[,-22]
#predictions <- predict(selected_model, newdata = test_data, type = "class")
predicted.classes <- modele.backward %>% predict(test2)
head(predicted.classes)
table_pred <- table(predicted.classes)
print(table_pred)
table(test$fetal_health)


```

```{r}

accuracyLG <- sum(as.character(predicted.classes) == as.character(test$fetal_health)) / length(predicted.classes)
accuracyLG

```

Decision Tree

```{r}
set.seed(123)
library(rpart)
library(rpart.plot)
Tree <- rpart(fetal_health ~., data = training, method = "class")
TC=Tree$cptable
which.min(TC[,4])# l'erreur minimale 
optcp=TC[which.min(TC[,4]),1]# optimal CP ( la valeur de CP qui correspond à l'erreur minimale)
opnsplit=TC[which.min(TC[,4]),2] # Nb de split(divisions) optimal 
opnsplit

#TOptimal <- rpart(survived~., data=ptitanic, cp=optcp, control=rpart.control(minsplit =opnsplit ))# arbre optimal
TOptimale<-rpart(fetal_health~.,data=training, cp=optcp, control=rpart.control(minsplit =opnsplit ))
rpart.plot(TOptimale)



```

```{r warning=FALSE}
#pred=predict(TOptimal,test,type ="vector")
#pred
#table(pred)# les classes prédites
pred=predict(TOptimale,test2, type = "class")
table(test$fetal_health)# les classes réelles
table_pred <- table(pred)
print(table_pred)

library(caret)
TC2 <- confusionMatrix(pred, test$fetal_health)
accuracy <- TC2$overall['Accuracy']

print(TC2$table)
print(accuracy)
#TC2=table(test$fetal_health,table_pred)# confusion matrix ( Y réel vs Y prédit )
#accuracy <- sum(pred == (test$fetal_health)) / length(pred)
#accuracy


```
Random Forest
```{r warning=FALSE}
set.seed(123)
library(randomForest)
modelRF <- randomForest(fetal_health ~ ., data = training)
predictForest = predict(modelRF, newdata = test2)
table(test$fetal_health, predictForest)

accuracyRF <- sum(as.character(predictForest) == as.character(test$fetal_health)) / length(predictForest)
accuracyRF
#library(caret)
#TC2 <- confusionMatrix(predictForest, test$fetal_health)
#accuracy <- TC2$overall['Accuracy']

#print(TC2$table)
#print(accuracy)

```
k-Means

```{r}
set.seed(123)
fetal_clust <- kmeans(training[,-22], centers = 3)
summary(fetal_clust)
cluster_assignments <-fetal_clust$cluster#Cluster number for each of the observations
head(cluster_assignments, 4)
actual_labels <- training$fetal_health
classification_accuracy <- sum(cluster_assignments == actual_labels) / length(actual_labels)
print(classification_accuracy)

```
Modele comparison

```{r}

all_accuracies <- as.data.frame(rbind(LG = accuracyLG, DT = accuracy, RF = accuracyRF, km=classification_accuracy ))
all_accuracies$model <- rownames(all_accuracies)
ggplot(all_accuracies, aes(x = model, y = Accuracy, fill = model)) +
  geom_bar(stat="identity") +
  ggtitle('Comparison of Overall Model Accuracy') +
  geom_text(aes(label = round(Accuracy,2)) , 
            colour = "white", size = 4,
            vjust = 1.5, position = position_dodge(.9))
```

